### 2016-04-27 techwriting/journal/ref_alg.tex
start journal writing

\subsection{Bit-level Variable Substitution (BLVS)}
\label{sec:blvs}
The remainder from \emph{Spoly} contains bit-level PS variables, and our objective is to get a polynomial contains only word-level PS variables
. One possible method is to rewrite bit-level variables in term of function on word-level
variables, i.e. $s_i = \mathcal{G}(S)$, then do substitution. A Gaussian-elimination-fashion approach could be applied to
compute corresponding $\mathcal{G}(S)$ efficiently.

@@

First, compute $f_0^2: s_0+s_1\alpha^2+S^2$. Apparently variable $s_0$ can be
eliminated by operation 
\begin{align}
f_1 =& f_0 + f_0^2: \nonumber\\
&(\alpha^2+\alpha)s_1+S^2+S\nonumber
\end{align}
Now we can solve univariate polynomial equation $f_1 = 0$ and get solution
$$s_1 = S^2 + S$$
Use this solution we can easily solve equation $f_0 = 0$. The result is
$$s_0 = \alpha S^2+(1+\alpha)S$$
\end{Example}
In this approach it is easy to get word-level variable representation for each bit-level PS variables. By substitution, a new polynomial in word-level PS/NS variables
could be attained.

>>>>

\subsection{Bit-level Variable Substitution (BLVS)}
\label{sec:blvs}
The remainder from \emph{Spoly} contains bit-level PS variables, and our objective is to get a polynomial contains only word-level PS variables
. One possible method is to rewrite bit-level variables in term of function on word-level
variables, i.e.
\begin{equation}
\label{eqn:BLVS}
s_i = \mathcal{G}(S)
\end{equation}
then do substitution. One way to obtain Eqn.\ref{eqn:BLVS} is to apply a Gaussian-elimination-fashion approach, which could 
compute corresponding $\mathcal{G}(S)$ efficiently with time complexity $O(k^3)$.

@@

First, compute $f_0^2: s_0+s_1\alpha^2+S^2$. Apparently variable $s_0$ can be
eliminated by operation 
\begin{align}
f_1 =& f_0 + f_0^2: \nonumber\\
&(\alpha^2+\alpha)s_1+S^2+S\nonumber
\end{align}
Now we can solve univariate polynomial equation $f_1 = 0$ and get solution
$$s_1 = S^2 + S$$
Use this solution we can easily solve equation $f_0 = 0$. The result is
$$s_0 = \alpha S^2+(1+\alpha)S$$
\end{Example}

Another formalized method is to solve following system of equations
\begin{align}
\label{eqn:betamat}
\begin{bmatrix}
S \\
S^2 \\
S^{2^2} \\
\vdots \\
S^{2^{k-1}}
\end{bmatrix}
&=
\begin{bmatrix}
\beta & \beta^{2} & \beta^{2^2} & \cdots & \beta^{2^{k-1}}\\
\beta^{2} & \beta^{2^2} & \beta^{2^3} & \cdots & \beta \\
\beta^{2^2} & \beta^{2^3} & \beta^{2^4} & \cdots & \beta^2\\
\vdots & \vdots & \vdots & \ddots & \vdots \\
\beta^{2^{k-1}} & \beta & \beta^2 & \cdots & \beta^{2^{k-2}}
\end{bmatrix}
\begin{bmatrix}
s_0\\
s_1\\
s_2\\
\vdots\\
s_{k-1}
\end{bmatrix}
\end{align}
Let ${\bf s}$ be a vector of $k$ unknowns $s_0,\dots,s_{k-1}$,
then Eqn.\ref{eqn:betamat} can be solved by using Cramer's rule:
\begin{equation}
\label{eqn:Cramer}
s_i = \frac{\Mi}{\M}, \ \ 0\leq i\leq k-1, \M \neq 0
\end{equation}
where ${\bf M_i}$ denotes a coefficient matrix replacing $i$-th column in ${\bf M}$ with
vector ${\bf S} = [S~~S^2~~\cdots~~S^{2^{k-1}}]^T$.

Notice that ${\bf M}$ is constructed by squaring a row and assigning it to next row,
therefore its determinant belongs to a special sort of determinants:
\begin{Definition}
Let $\{\alpha_0,\alpha_1,\dots,\alpha_{k-1}\}$ be a set of $k$ elements of $\Fpk$. Then the determinant
\begin{align}
\det{M(\alpha_0,\dots,\alpha_{k-1})} &= 
\begin{vmatrix}
\alpha_0 & \alpha_1 & \cdots & \alpha_{k-1} \\
\alpha_0^p & \alpha_1^p & \cdots & \alpha_{k-1}^p \\
\vdots & \vdots & \ddots & \vdots \\
\alpha_0^{p^{k-1}} & \alpha_1^{p^{k-1}} & \cdots & \alpha_{k-1}^{p^{k-1}}
\end{vmatrix}
\end{align}
is called the {\bf Moore determinant} of set $\{\alpha_0,\dots,\alpha_{k-1}\}$.
\end{Definition}
Moore determinant can be written as an explicit expression
\begin{equation}
\label{eqn:Moore}
\det{M(\alpha_0,\dots,\alpha_{k-1})} = \alpha_0 \prod_{i=1}^{k-1} \prod_{c_0,\dots,c_{i-1}\in \Fp} (\alpha_{i} - \sum_{j=0}^{i-1}c_j\alpha_j)
\end{equation}
We use an example to help understanding the notations in Eqn.\ref{eqn:Moore}:
\begin{Example}
Let $\{\alpha_0,\alpha_1,\alpha_2\}$ be a set of elements of $\mathbb F_{2^3}$. Then
\begin{align}
\det{M(\alpha_0,\alpha_1,\alpha_2)} &= 
\begin{vmatrix}
\alpha_0 & \alpha_1 & \alpha_2 \\
\alpha_0^2 & \alpha_1^2 & \alpha_2^2 \\
\alpha_0^4 & \alpha_1^4 & \alpha_2^4
\end{vmatrix} \\ \nonumber
=& \alpha_0 \prod_{i=1}^{2} \prod_{c_0,\dots,c_{i-1}\in \Ftwo} (\alpha_{i} - \displaystyle\sum_{j=0}^{i-1}c_j\alpha_j)
\end{align}
First, let $i=1$,  we obtain $c_0\in \Ftwo$. When $c_0=0$, the product term equals to $\alpha_1$; when $c_0=1$ it equals to $(\alpha_1-\alpha_0)$.
Then let $i=2$, we obtain ${c_0,c_1}\in \Ftwo$, they can take value from $\{0,0\},\{0,1\},\{1,0\}$ and $\{1,1\}$.
We add 4 more product terms $\alpha_2,(\alpha_2-\alpha_1),(\alpha_2-\alpha_0),(\alpha_2-\alpha_0-\alpha_1)$ respectively.

Thus, the result is
\begin{equation}
\det{M(\alpha_0,\alpha_1,\alpha_2)} = \alpha_0\alpha_1(\alpha_1-\alpha_0)\alpha_2(\alpha_2-\alpha_1)(\alpha_2-\alpha_0)(\alpha_2-\alpha_0-\alpha_1)
\end{equation}
\end{Example}
We discover through investigation that $\M$ has a special property when the set of elements forms a basis. The proof is given below:
\begin{Lemma}
\label{lemma:Moore}
Let $\{\alpha_0,\alpha_1,\dots,\alpha_{k-1}\}$ be a normal basis of $\Fpk$ over $\Fp$. Then
\begin{equation}
\det{M(\alpha_0,\alpha_1,\dots,\alpha_{k-1})} = 1
\end{equation}
\end{Lemma}
\begin{proof}
According to the definition Eqn.\ref{eqn:Moore}, the Moore determinant consists of all possible linear combinations of 
$\{\alpha_0,\alpha_1,\dots,\alpha_{k-1}\}$ with coefficients over $\Fp$. If $\{\alpha_0,\alpha_1,\dots,\alpha_{k-1}\}$
is a (normal) basis, then all product terms are distinct and represents all elements in the field $\Fpk$. Since the product 
of all elements of a field equals to 1, the Moore determinant $\M=1$.
\end{proof}
Applying Lemma \ref{lemma:Moore} to Eqn.\ref{eqn:Cramer} gives
\begin{equation}
s_i = \Mi,~~0\leq i\leq k-1
\end{equation}
where $\Mi$ can be easily computed using Laplace expansion method, with complexity $O(k!)$.

In this approach it is easy to get word-level variable representation for each bit-level PS variables. 
By substitution, a new polynomial in word-level PS/NS variables could be attained.
